---
title: "TableContainer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TableContainer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(TableContainer)
```


# TableContainer

`TableContainer` is an R package that provides a lightweight and flexible container for managing tabular data with associated row and column annotations. It is inspired by Bioconductor's `SummarizedExperiment` but does not rely on Bioconductor dependencies, making it easier to integrate into various workflows.

## Features

- **Matrix-like Data Storage**: Store data in a matrix or data frame format.
- **Row and Column Annotations**: Add metadata to rows and columns using data frames.
- **Flexible Metadata**: Attach arbitrary metadata to the container.
- **Automatic Annotation Updates**: Automatically update annotations when subsetting the table.


## Creating a TableContainer

You can create a `TableContainer` object using the `TableContainer()` constructor:

```{r}
# Example data
tbl <- matrix(1:12, nrow = 3, ncol = 4)
row_dt <- data.frame(row1 = 1:3, row2 = letters[1:3])
col_dt <- data.frame(col1 = 1:4, col2 = letters[1:4])
metadata <- list(meta1 = "meta1", meta2 = "meta2")

# Create a TableContainer
container <- TableContainer(
  table = tbl,
  rowData = row_dt,
  colData = col_dt,
  metadata = metadata
)

# Check the container
container
```

## Subsetting

Subset the `TableContainer` object while maintaining consistency in annotations:

### Subset rows and columns
```{r}
subset <- container[1:2, 2:4]
subset
```

The row and column annotations are automatically updated to reflect the subsetted data.
```{r}
rowData(subset)
colData(subset)
```

### Subset a single row
```{r}
container[1, ]
```
This is equivalent to:
```{r}
container[1]
```

### Subset a single column
```{r}
container[, 2]
```

### Subset a single cell
```{r}
container[1, 2]
```


## Accessing and Modifying Data

You can access and modify the table, row annotations, column annotations, and metadata using accessor methods:

```{r}
# Access data
tblData(container)
rowData(container)
colData(container)
metadata(container)

# Modify data
tblData(container) <- matrix(13:24, nrow = 3, ncol = 4)
rowData(container) <- data.frame(newRow = 1:3)
colData(container) <- data.frame(newCol = 1:4)
metadata(container) <- list(newMeta = "updated metadata")

## check the updated container
container
```

